### 2016-12-15 paper/PacketSamplingFirewalls.tex
Answered a bunch of questions in notes/, started moving paper and proposal to LaTeX

\documentclass[fleqn]{article}%
\usepackage{amsmath}%
\usepackage{amsfonts}%
\usepackage{amssymb}%
\usepackage{graphicx}
%-------------------------------------------
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}[theorem]{Acknowledgement}
\newtheorem{algorithm}[theorem]{Algorithm}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{case}[theorem]{Case}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{solution}[theorem]{Solution}
\newtheorem{summary}[theorem]{Summary}
\newenvironment{proof}[1][Proof]{\textbf{#1.} }{\ \rule{0.5em}{0.5em}}
\setlength{\textwidth}{7.0in}
\setlength{\oddsidemargin}{-0.35in}
\setlength{\topmargin}{-0.75in}
\setlength{\textheight}{9.2in}
\setlength{\parindent}{0.0in}
\begin{document}

\begin{center}
  \textbf{Packet Sampling Firewalls \\ Austin Voecks}\\
\end{center}

\textbf{Outline}
\begin{itemize}
    \item Problem
    \item Abstract
    \item Subsampling
    \item Randomized Subsampling
    \item Dynamic Subsampling
    \item Heuristic Based Rules
    \item IPFW
    \item Benchmarking Performance
    \item Results
    \item Security Implications
    \item Related Work
    \item Conclusion
    \item References
\end{itemize}

\textbf{Problem} \\
Provide non-disruptive backend network monitoring to OneFS at 40Gb speeds.

\textbf{Abstract} \\
The performance of FreeBSD’s IPFW can be increased to support 40Gb network
interfaces without additional specialized hardware through dynamic
probabilistic rule matching. The probability of rule matching will be based on
current system load, allowing it to adapt to changing demands on the system.
The efficacy of packet stream subsampling will also be explored.

\textbf{Subsampling} \\
Ideally, all network traffic would be carefully inspected; this is impractical
due to the limitations of CPU and bus speeds. When logging, it’s also limited
by disk speed. Subsampling traffic provides a way take the temperature of the
network and increase chances of detecting abnormal behavior; some monitoring is
better than none. To provide full packet inspection to every packet on system,
the network links would have to be split into 10Gb links, each passed through
dedicated hardware.

\end{document}